The complex network of trophic interactions in a subAntarctic Marine Protected

### 2023-03-28 MS_ToSubmit/MS_Burdwood_foodweb.Rmd
Add LS & IR minor comments

---
title: The complex network of trophic interactions in a subAntarctic Marine Protected
  Area
author:
  - name: Tomás I. Marina^a^
    email: tomasimarina@gmail.com
  - name: Irene R. Schloss^a^^,^^b^
  - name: Luciana Riccialdelli^a^
address:
  - code: 1
    address: Centro Austral de Investigaciones Cientificas (CADIC-CONICET), Argentina
  - code: 2
    address: Instituto Antartico Argentino (IAA), Argentina
abstract: |
  The abstract goes here.

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as well as the urgent need for ocean protection have driven an unprecedented increase in the number of MPAs worldwide in recent years [@Roberts2017; @Sala2018]. Globally, the total area of the world ocean designated under marine protection adds up to nearly 29,600,000 km^2^, distributed across nearly 18,444 MPAs and covering 8.16% of the ocean’s surface [@IUCN2023], and therefore approaching the 10% goal of the Convention of Biological Diversity [@CBD2004]. Despite this progress, recent reports have shown that effective protection has been overestimated because it includes areas that are not yet effectively protected (only declared) as well as areas that allow significant extractive activities [@Sala2018].

In the subAntarctic region, the level of protection is mainly associated to oceanic islands, such as South Georgia and the South Sandwich, Bouvet, Prince Edward, and Macquarie [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Banco Burdwood I and II (MPA N-BB), which forms the focus of this work, is unique since they are associated to a submarine plateau and the southern adjacent deep slope region, respectively [@Falabella2017]. Added to this, such MPAs are part of a network of protected areas in the subAntarctic (jointly with MPA Yaganes), that contributes to protect and strengthen the functioning of this southern region to maintain global ocean health.

Many of these MPAs centre on the presence of particularly vulnerable, keystone, or charismatic species; the presence of large numbers (or proportions) of endemic species; and the presence of high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a sensitive and biodiverse benthic habitat that was barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by a high biomass of vulnerable and fragile species (mainly Porifera and Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResources2009]. Also, it provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge, or nursery area [@GarciaAlonso2018; @Delpiani2020; @Troccoli2020]. The maintenance of this singular community is related to the oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients coming from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time for the area [@APN2022].

Recently, the identification of the most influential species has become a crucial task for conservation purposes in local and regional contexts to maintain ecosystem integrity and the supply of ecosystem services for the southwestern South Atlantic Ocean [@Franco2020; @Bergagna2020; @Riccialdelli2020].
In this regard, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem is regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure than other subAntarctic areas, like the Atlantic coast of Tierra del Fuego. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are characterised by an unexpected complexity in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @APN2022], there is a need for a better understanding of species interactions' complexity and structure.
This aspect can be tackled by analysing one of the most-frequent relationships occurring between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, which represents the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and respond to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Here it is worth mentioning that anthropogenic threats have been recently reported for the MPA N-BB region [@Cossi2021; @Fioramonti2022].

In the present work we present the first, detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this we applied a network approach to a highly-resolved food web. The objective was twofold: characterise the food web in terms of complexity and structure, and identify the species' role in such network framework.

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws 26875 and 27490, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, respectively [@APN2023]. They are located 150 km east of Isla de los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1).
The MPA Namuncurá - Banco Burdwood I comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC overall [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. Such deep slope form the basis of the MPA Namuncurá - Banco Burdwood II that protects about 32,000 km^2^. Intense upwelling and mixing occur over it, entraining deep nutrient rich waters into the photic layer [@Piola2009; @Matano2019], and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

Given the evidence collected during several research cruises about the oceanographic and ecological processes connecting MPAs Namuncurá - Banco Burdwood I and II [references in @APN2022], a joint management plan was recently proposed [@APN2022]. Three categories were defined according to the authorized activities: 1) Marine National Reserve (MNR), an area with the objective of conserving marine biodiversity, landscape quality and large-scale ecological processes, allowing scientific, educational, recreational uses and the sustainable use of one or more of its natural resources; 2) Marine National Park, an area with the objective of conserving marine biodiversity, landscape quality and large-scale ecological processes (MNP); and 3) Restricted Marine National Reserve (RMNR), an area of maximum permanent or temporary protection with the objective to preserve marine biodiversity and ecological processes in the long term, minimizing anthropogenic impact, and allowing scientific research and monitoring, as well as control and surveillance activities. Figure 1 shows the geographic location of these categories within MPAs Namuncurá - Banco Burdwood I and II.

```{r figure1, echo=FALSE, fig.width = 5, fig.height = 10, out.width = '100%', fig.cap="Map of the Marine Protected Areas Namuncurá - Banco Burdwood I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve."}

@@

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases we followed it when specific data on species, in the taxonomic sense, were not available. In some cases we collapsed species when taxa shared the same set of predators and prey (trophic similarity, @Martinez1991), one of the aggregation methods that better preserves food web functional properties [@Gauzens2013]. Details about this can be found in Supplementary Material (Table S2).

With the gathered trophic data we constructed an interaction matrix of pairwise interactions; a value of 1 or 0 was assigned to each element a_ij of the matrix depending on whether the j-species preyed or not on the i-species. Then we transformed such matrix into an oriented graph with L trophic interactions between S nodes or species. The orientation or direction of the graph follows the flow of energy and matter in the network, from prey to predator.

@@

We analysed the MPA Namuncurá - Banco Burdwood network of trophic interactions, hereafter food web, at two levels: A) network, considering species and interactions of the whole network; and B) species, considering interactions and species related to a particular species (Table 1).

The network-level analysis aims to characterise the network of trophic interactions in terms of complexity and structure. For this we calculated several network properties commonly used to describe empirical food webs [@Pascual2005]: (1) number of species S; (2) number of interactions or links L; (3) link density L/S; (4) Connectance L/S^2; (5) omnivory Omn; and (6) small-world pattern. In order to explore the small-world phenomenon we analysed the properties of interest: characteristic path length (CPL) and clustering coefficient (CC). The CPL is defined as the average shortest path length between all pairs of nodes [@Watts1998]. Here, CPL was calculated as the average number of nodes in the shortest path $CPL_{Min} (i,j)$ between all pairs of nodes $S(i,j)$ in a network averaged over $S(S-1)/2$ nodes:

@@

$$
CC_i = \frac{2E_i}{K_i(K_i-1)}
$$
where $E_i$ is the effective number of interactions between $K_i$ nearest-neighbor nodes of node $i$ and the maximal possible number of such interactions [@Newman2003].
With the aim of testing whether the food web presented the small-world pattern, we compared the empirical values of CPL and CC with those resulted from 1000 randomly generated networks with the same size (S) and number of interactions (L), following the method proposed by @Marina2018a.

@@

The species-level analysis aims to describe the species' role in the food web. For this we considered the following properties: (1) betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4) topological role TR; and (5) trophic level TL (Table 1). Topological roles refers to the fact that food webs tend to naturally organize in non-random, modular patterns [@Grilli2016]. Species can play different roles in this respect, according to the pattern of interactions they have within their own module and/or across modules. We computed the topological role for each species, classified as: module hub, species with a relatively high number of interactions, but most within its own module; module specialist, species with relatively few interactions and most within its own module; module connector, species with relatively few interactions mainly between modules; and network connector, species with high connectivity between and within modules [@Guimera2005].

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web consisted of 1778 predator-prey interactions and 379 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density of 4.69, meaning the average number of interactions per species, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient was higher than the random networks (Table 2).

@@

```{r figure2, echo=FALSE, fig.width = 5.5, fig.height = 10, fig.align='center', fig.cap="Graph of the food web for the MPA Namuncurá - Banco Burdwood. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicate the trophic level."}
# Food web graph

@@

The majority of the species of the food web were consumers, 336 out of 379; the rest were primary producers, such as diatoms (phytoplankton), and non-living food sources like detritus and necromass.

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). The most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight high-trophic level species with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Phalacrocorax_atriceps* and *Eudyptes chrysocome* (sea birds), and *Lagenorhynchus cruciger* and *Mesoplodon bowdoini* (marine mammals) (Table S3).

We also found a negative significant relationship with closeness, however less evident, meaning that low-TL species are relatively closer to any other species in the food web (Figure 4B). Species of genera *Calanus* and *Euphausia*, and species of Brachiopoda, all of them with TL < 3, registered the highest values of closeness (Table S3).

@@

Taking into account the topological role, 'module specialist' species were the most frequent and presented a wide TL range (1 - 4.77); 'module hub' was constrained to mid-TL species (2.48 - 3.92); 'module connector' from low to mid-TLs (1 - 3.86); and 'network connector', the least frequent, had all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D, see Figure S2 for species' topological roles in a food web graph framework). Here it is important to highlight the two latter topological roles, because they are responsible for linking modules and maintaining the connectivity of the food web: 40 species (5 network connectors + 35 module connectors) from 20 different functional groups with a TL range = 1 - 3.86. The following were the 5 species with a network connector role: Detritus, Diatoms (benthic, centric and pennate), and *Patagonotothen ramsayi*. The 35 species with a module connector role represent these functional groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, Detritus, Diatoms, Dinoflagellates, Echinodermata, Fish (bentho-pelagic, demersal, chondrichthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, Sea birds, and Zooplankton (see Supplementary Material Table S3 for the identity of the species).

@@

The food web of the MPA Namuncurá - Banco Burdwood ecosystem, analysed in this study, is one of the most highly-resolved networks of trophic interactions ever studied not only for a high-latitude marine ecosystem but also for a marine protected area worldwide. It is of paramount importance to consider the complexity of species interactions in order to gain insights into the structure and functioning of the ecosystem, since the aggregation of species might mask food web properties and/or produce type II errors (false positive) [@Martinez1993; @Gauzens2013].

Food web connectance is a feature that resumes the complexity of the network, but more importantly it is an emergent property of pairwise species interactions [@Poisot2014]. It contains information regarding how interactions within an ecological network are distributed, and predicts reasonably well key dynamical properties of ecological networks [@Dunne2002a].
Complex marine food webs (i.e. with more than 25 trophic species) show connectance values ranging from 0.01 - 0.27 [@Marina2018a]. In particular, food webs from high-latitude regions tend to exhibit a connectance closer to the minimum (between 0.01 and 0.05) [@Santana2013; @Kortsch2015; @Rodriguez2022]. Whether food webs display a low or a high connectance helps to better comprehend ecosystem's synthetic properties like robustness. In this sense, empirical analyses support the notion that highly-connected ecological networks are robust against external perturbations such as the introduction of new (e.g., invasive) species [@Smith-Ramesh2017] as well as species removal (e.g., local extinction) [@Dunne2002; @Montoya2003]. The connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01) is one of the lowest reported so far for these regions; five times lower than that of Beagle Channel, an adjacent area [@Rodriguez2022].

The degree distribution or the distribution of number of interactions per species is the core for the structure of species interactions, which influences the opportunities for multiple species to persist in the long term and, therefore, their coexistence [@Godoy2018]. The food web for the MPA N-BB presents an asymmetric degree distribution. This is a pattern that we have identified at different levels of analysis: food web, predator, prey, and functional group. Such asymmetry is a well-known feature in empirical complex food webs in particular [@Dunne2002a; @Montoya2003; @Stouffer2005], and has received great attention in complex networks in general [@Albert2002; @Newman2003].
The degree distribution affects the resilience of complex food webs against random failures and intentional attacks: right-skewed distributions, like the one described in this study, are more vulnerable to the removal of the most-connected species or hubs with the potential of producing secondary extinctions and a catastrophic fragmentation of the network [@Albert2000; @Dunne2002; @Eklof2006].

It is suggested that the small-world pattern is not frequent in complex marine food webs, mainly due to a low clustering coefficient compared to random networks [@Dunne2002b; @Marina2018a]. However, the food web of the MPA N-BB does display a small-world pattern, characterizing a network with short path length and high clustering coefficient. Consequences of the small-world pattern in food webs are of great importance in recognizing evolutionary paths and the vulnerability to perturbations [@Montoya2002]. On the one hand, a short path length implies a rapid spread of an impact (e.g., contaminant, population fluctuation, local extinction) throughout the network and, at the same time, a more potentially adaptive dynamics in face of external perturbations [@Montoya2002; @Williams2002]. On the other hand, a high clustering coefficient indicates the formation of subnetworks composed only by the neighbours of particular species. This translates into a greater resistance of the network due to the confinement of perturbations mainly within subnetworks and not between them [@Kortsch2019; @Heer2020]. Overall, a small-world topology provides ecological networks with a greater resilience and resistance [@Bornatowski2017; @Dormann2017].

Omnivory acts as a buffer to changes as the ecosystem presents alternative energy pathways in the face of perturbations; i.e., reducing the risk of cascading extinctions following primary loss of species [@Borrvall2000]. This is based in the fact that omnivores are species able to adapt faster and at a wider range of environmental conditions by changing their foraging habits to feed on the most abundant prey [@Fagan1997]. Furthermore, omnivory can be analysed from the interaction point of view: theoretical studies have identified omnivorous interactions as a possible candidate for a keystone interaction, sensu @Kadoya2018, highlighting the importance on stabilizing food web dynamics [@McCann1997; @Neutel2002]. The high proportion of omnivory in the food web of the MPA N-BB suggests that the network might be robust to variations in prey abundances, which could increase food web's persistence and stability [@Stouffer2010].

To resume, the food web of the MPA Namuncurá - Banco Burdwood presents a combination of network properties that make it unique in terms of network resolution, complexity, and structural pattern. All this suggest that the ecosystem might be fragile to external perturbations targeting high connected species. However, structural properties might provide resilience and resistance with the final outcome of a rearrange structure maintaining their functions.

@@

Describing species’ roles in food webs provides a toolbox with which to assess the significance of species in terms of community's functioning and overall stability [@Thebault2010; @Cirtwill2018]. We employed a range of descriptors to characterise the dynamic and multifaceted nature of the species comprising the MPA N-BB food web.

>>>>

---
title: The complex network of trophic interactions in a subAntarctic Marine Protected
  Area
author:
  - name: Tomás I. Marina^a^
    email: tomasimarina@gmail.com
  - name: Irene R. Schloss^a^^,^^b^
  - name: Luciana Riccialdelli^a^
  - name: To be continued(?)^c^
address:
  - code: 1
    address: Centro Austral de Investigaciones Cientificas (CADIC-CONICET), Argentina
  - code: 2
    address: Instituto Antartico Argentino (IAA), Argentina
  - code: 3
    address: INIDEP, UBA-CONICET, IIMyC-CONICET, MACN-CONICET, UNC-CONICET
abstract: |
  The abstract goes here.

@@

The evidence of benefits provided by Marine Protected Areas (MPAs) as well as the urgent need for ocean protection have driven an unprecedented increase in the number of MPAs worldwide in recent years [@Roberts2017; @Sala2018]. Globally, the total area of the world ocean designated under marine protection adds up to nearly 29,600,000 km^2^, distributed across nearly 18,444 MPAs and covering 8.16% of the ocean’s surface [@IUCN2023], and therefore approaching the 10% goal of the Convention of Biological Diversity [@CBD2004]. Despite this progress, recent reports have shown that actual protection has been overestimated because it includes areas that are not yet effectively protected (only declared) as well as areas that allow significant extractive activities [@Sala2018].

In the sub-Antarctic region, the level of ocean protection is mainly associated to oceanic islands, such as the South Georgia and South Sandwich, Bouvet, Prince Edward, and Macquarie islands [@IUCN2023]. Interestingly, the case of the MPAs Namuncurá - Banco Burdwood I and II (MPA N-BB), which is the focus of this work, is unique since these MPAs are associated to a submarine plateau and its southern adjacent deep slope region, respectively [@Falabella2017]. Added to this, such MPAs are part of a network of protected areas in the sub-Antarctic (jointly with MPA Yaganes), that aims to protect and strengthen the functioning of this southern region to maintain global ocean health.

Many of these MPAs focus on the presence of particularly vulnerable, keystone, or charismatic species, the presence of large numbers (or proportions) of endemic species, and/or the presence of high biodiversity across taxonomic levels [@Hogg2016]. Indeed, the MPA N-BB was created to protect a sensitive and biodiverse benthic habitat that was barely known [@Schejter2016; @Falabella2017]. The benthic community is featured by a high biomass of vulnerable and fragile species (mainly Poriferaand Cnidaria) that meet the characteristics of vulnerable marine ecosystems [@Schejter2021], here defined as sites that present taxa with a minimum biomass of 10 kg per 1200 m^2^ [@CommissionfortheConservationofAntarcticMarineLivingResources2009]. Also, it provides habitat to several small-sized species [@Lopez-Gappa2018; @MartinSirito2019; @Schejter2019], and has an important role in the life history of fishes as a food source, refuge and nursery area [@GarciaAlonso2018; @Delpiani2020; @Troccoli2020]. The maintenance of this singular community is related to local and regional oceanographic processes, including the circulation of the rich Malvinas (Falkland) current in the area [@Piola1989; @Guerrero1999] and the upwelling and mixing phenomena [@Matano2019]. The input of nutrients coming from the Malvinas (Falkland) current also supports a diverse plankton community [@Guinder2020]. Overall, 811 benthic and plankton species have been identified for the MPA N-BB ecosystem, where 349 were reported for the first time for the area [@APN2022]. However, identifying the main species for maintenance of ecosystem services and health as well as for management and conservation is essential.

Recently, the structure of the southwestern South Atlantic Ocean has been proposed to be under a 'wasp-waist' control, meaning that the structure and dynamics of the ecosystem is regulated primarily by mid-trophic level species (e.g., fishes, crustaceans) [@Padovani2012; @Saporiti2015; @Riccialdelli2020]. In particular, the ecosystem of the MPA N-BB shows a more pronounced 'wasp-waist' structure than other sub-Antarctic areas, such as the Atlantic coast of Tierra del Fuego. The Fuegian sprat *Sprattus fuegensis* and longtail southern cod *Patagonotothen ramsayi* are considered the most plausible 'wasp-waist' species [@Riccialdelli2020].

High-latitude marine ecosystems, such as the MPA N-BB, are characterized by an unexpected complexity in terms of biodiversity and ecological interactions [@Day2013; @Kortsch2019; @Cordone2020; @Trathan2021]. Although there is a robust knowledge about the complexity considering the richness of the benthic and plankton communities in the MPA N-BB ecosystem [@Schejter2016; @Guinder2020; @Schejter2020; @APN2022], there is a need for a better understanding of species interactions' complexity and structure.
This aspect can be tackled by analyzing one of the most-frequent relationships occurring between species: the predator-prey interaction [@Bascompte2009]. The sum of predator-prey or trophic interactions of a particular region is referred to as a food web, which represents the roadmap for matter and energy flow in an ecosystem.
In recent years, network approaches have been successfully applied to study complex high-latitude marine ecosystems, improving our knowledge on structure, functioning, and response to environmental/anthropogenic changes [@Kortsch2015; @Cordone2018; @Funes2022; @Marina2023]. Among anthropogenic threats, it is worth mentioning that contaminants like mercury and microplastics have been recently reported as important threats for the MPA N-BB region [@Cossi2021; @DiMauro2022; @Fioramonti2022].

In the present work, we present the first, detailed analysis of the network of predator-prey interactions, hereafter food web, for the MPA N-BB ecosystem. For this, we applied a network approach to a highly resolved food web. The objective was twofold: characterize the food web in terms of complexity and structure, and identify the species' role in such network framework.

@@

The MPAs Namuncurá - Banco Burdwood I and II, created by National Laws 26.875 and 27.490, comprise a shallow submarine plateau called Burdwood Bank (BB) and a deep slope that reaches 4000 m in depth, N-BB I and N-BB II, respectively [@APN2023]. They are located 150 km east of isla de los Estados and 200 km south from Malvinas/Falkland Islands (Figure 1).
The MPA N-BB I comprises nearly 34,000 km^2^ circumscribed by the 200 m isobath, between 54º–55ºS and 56º–62ºW, with a slight slope extended nearly 370 km east–west. Physical features in the BB are fairly stable, with salinity averaging 34 all year round and temperature ranging between 4 and 8ºC [@Guerrero1999; @Acha2004; @Piola2009]. The BB is surrounded by steep flanks of up to 4000 m depth through which strong currents circulate [@Piola1989; @Reta2014; @Matano2019]. Such deep slope form the basis of N-BB II that protects about 32,000 km^2^. Intense upwelling and mixing occur over it, entraining deep nutrient-rich waters into the photic layer [@Piola2009; @Matano2019], and resulting in a fairly homogeneous water column both spatially and temporally [@Glorioso1995; @Guerrero1999; @Matano2019].

Given the evidence collected during several research cruises about the oceanographic and ecological processes connecting MPAs N-BB I and II [references in @APN2022], a joint management plan was recently proposed [@APN2022].

```{r figure1, echo=FALSE, fig.width = 5, fig.height = 10, out.width = '100%', fig.cap="Marine Protected Areas Namuncurá - Banco Burdwood I (MNR and MNP, northern section) and II (MNR and SMNR, southern section). Acronyms indicate categories according to the management plan: MNR - Marine National Reserve, MNP - Marine National Park and RMNR - Restricted Marine National Reserve."}

@@

Due to a lack of trophic data resolution for some species inhabiting the study area, we followed the concept of trophic species, here defined as aggregated groups of taxa. In most cases, we followed it when specific data on species, in the taxonomic sense, were not available. In some cases, we collapsed species when taxa shared the same set of predators and prey (trophic similarity, @Martinez1991), one of the aggregation methods that better preserves food web functional properties [@Gauzens2013]. Details about this can be found in Supplementary Material (Table S2).

With the gathered trophic data we constructed a matrix of pairwise interactions; a value of 1 or 0 was assigned to each element a_ij of the matrix depending on whether the j-species preyed or not on the i-species. Then we transformed such matrix into an oriented graph with L trophic interactions between S nodes or species. The orientation or direction of the graph follows the flow of energy and matter in the network, from prey to predator.

@@

We analyzed the MPA N-BB network of trophic interactions, or food web, at two levels: A) network, considering species and interactions of the whole network; and B) species, considering interactions and species related to a particular species (Table 1).

The network-level analysis aims to characterizee the food web in terms of complexity and structure. For this we calculated several network properties commonly used to describe empirical food webs [@Pascual2005]: (1) number of species S; (2) number of interactions or links L; (3) link density L/S; (4) connectance L/S^2; (5) omnivory Omn; and (6) small-world pattern. In order to explore the small-world phenomenon we analyzed the  characteristic path length (CPL) and the clustering coefficient (CC). The CPL is defined as the average shortest path length between all pairs of nodes [@Watts1998]. Here, CPL was calculated as the average number of nodes in the shortest path $CPL_{Min} (i,j)$ between all pairs of nodes $S(i,j)$ in a network averaged over $S(S-1)/2$ nodes:

@@

$$
CC_i = \frac{2E_i}{K_i(K_i-1)}
$$
where $E_i$ is the effective number of interactions between $K_i$ nearest-neighbor nodes of node $i$ and the maximal possible number of such interactions [@Newman2003].
With the aim of testing whether the food web presented the small-world pattern, we compared the empirical values of CPL and CC with those resulting from 1000 randomly generated networks with the same size (S) and number of interactions (L), following the method proposed by @Marina2018a.

@@

The species-level analysis aims to describe the species' role in the food web. For this, we considered the following properties: (1) betweenness Btw; (2) closeness Cl; (3) trophic similarity TS; (4) topological role TR; and (5) trophic level TL (Table 1). Topological roles refers to the fact that food webs tend to naturally organize in non-random, modular patterns, where modules are defined as a group of species that interact more frequently among themselves than with species that are not members of the module [@Guimera2005]. Species can play different roles in this respect, according to the pattern of interactions they have within their own module and/or across modules. We computed the topological role for each species, classified as: module hub, species with a relatively high number of interactions, but most within its own module; module specialist, species with relatively few interactions and most within its own module; module connector, species with relatively few interactions mainly between modules; and network connector, species with high connectivity between and within modules [@Guimera2005].

@@

In terms of complexity, the MPA Namuncurá - Banco Burdwood food web (N-BB I and II) consisted of 1778 predator-prey interactions and 379 species, where 93% of them were defined at the species taxonomical level (Figure 2, Table S2). The food web presented a link density (e.g., the average number of interactions per species) of 4.69, and a connectance of 0.01. Almost half of the consumers were omnivores (0.48), feeding on sources at different trophic levels. The food web displayed a small-world pattern, meaning that the path length was lower and the clustering coefficient higher than the random networks (Table 2).

@@

```{r figure2, echo=FALSE, fig.width = 5.5, fig.height = 10, fig.align='center', fig.cap="Graph of the food web for the MPA Namuncurá - Banco Burdwood. Circles represent species and arrows trophic interactions. Circle diameter is relative to the number of interactions. Colour gradient indicates the trophic level."}
# Food web graph

@@

The majority of the species in the food web were consumers, 336 out of 379; the rest were primary producers, such as diatoms (phytoplankton), and non-living food sources like detritus and necromass.

We found different relationships between the species trophic level (TL) and the rest of the analysed species-level properties (Figure 4A-D). The most evident significant relationship was with trophic similarity, i.e. the higher the species' TL, the lower the trophic similarity or the higher the uniqueness in terms of trophic role (Figure 4C). Here it is noteworthy to highlight those high-trophic level species with low values of trophic similarity: *Bathyraja macloviana* and *Squalus acanthias* (Chondrichthyans), *Diplopteraster clarki* and *Pteraster sp* (echinoderms), *Phalacrocorax_atriceps* and *Eudyptes chrysocome* (sea birds), and *Lagenorhynchus cruciger* and *Mesoplodon bowdoini* (marine mammals) (Table S3).

We also found a negative significant relationship between TL and closeness, however less evident, meaning that low-TL species are relatively closer to any other species in the food web (Figure 4B). Species of genera *Calanus* and *Euphausia*, and species of Brachiopoda, all of them with TL < 3, registered the highest values of closeness (Table S3).

@@

Considering the topological role, 'module specialist' species were the most frequent and presented a wide TL range (1 - 4.77); 'module hub' was constrained to mid-TL species (2.48 - 3.92); 'module connector' from low to mid-TLs (1 - 3.86); and 'network connector', the least frequent, had all of its species in TL = 1, except for one with TL = 3.47 (Figure 4D, see Figure S2 for species' topological roles in a food web graph framework). Here it is important to highlight the two latter topological roles, because they are responsible for linking modules and maintaining the connectivity of the food web: 40 species (5 network connectors + 35 module connectors) from 20 different functional groups with a TL range = 1 - 3.86. The following were the five species with a network connector role: Detritus, Diatoms (benthic, centric and pennate), and *Patagonotothen ramsayi*. The 35 species with a module connector role represent these functional groups: Amphipoda, Bivalvia, Brachiopoda, Bryozoa, Cnidaria, Cumacea, Decapoda, Detritus, Diatoms, Dinoflagellates, Echinodermata, Fish (bentho-pelagic, demersal, chondrichthyes), Foraminifera, Polychaeta, Porifera, Pycnogonida, Sea birds, and Zooplankton (see Supplementary Material Table S3 for the identity of the species).

@@

The food web of the MPA Namuncurá - Banco Burdwood ecosystem analysed in this study is one of the most highly-resolved networks of trophic interactions ever studied not only for a high-latitude marine ecosystem but also for a marine protected area worldwide. It is of paramount importance to consider the complexity of species interactions in order to gain insights into the structure and functioning of the ecosystem, since the aggregation of species might mask food web properties and/or produce type II errors (false positive) [@Martinez1993; @Gauzens2013].

Food web connectance is a feature that resumes the complexity of the network, but more importantly, it is an emergent property of pairwise species interactions [@Poisot2014]. It contains information regarding how interactions within an ecological network are distributed, and predicts reasonably well key dynamical properties of ecological networks [@Dunne2002a]. Complex marine food webs (i.e. with more than 25 trophic species) show connectance values ranging from 0.01 - 0.27 [@Marina2018a]. In particular, food webs from high-latitude regions tend to exhibit a connectance closer to the minimum (between 0.01 and 0.05) [@Santana2013; @Kortsch2015; @Rodriguez2022]. Whether food webs display a low or a high connectance helps to better comprehend ecosystem's synthetic properties like robustness. In this sense, empirical analyses support the notion that highly-connected ecological networks are robust against external perturbations such as the introduction of new (e.g., invasive) species [@Smith-Ramesh2017] as well as species removal (e.g., local extinction) [@Dunne2002; @Montoya2003]. The connectance of the food web of the MPA Namuncurá - Banco Burdwood (0.01) is one of the lowest reported so far for these regions; in particular, it appears to be five times lower than that of Beagle Channel, an adjacent area [@Rodriguez2022].

The degree distribution or the distribution of number of interactions per species is the core for the structure of species interactions, which influences the opportunities for multiple species to persist in the long term and, therefore, their coexistence [@Godoy2018]. The food web for the MPA N-BB presents an asymmetric degree distribution. This pattern was identified at different levels of analysis: food web, predator, prey, and functional group. Such asymmetry is a well-known feature in empirical complex food webs in particular [@Dunne2002a; @Montoya2003; @Stouffer2005], and has received great attention in complex networks in general [@Albert2002; @Newman2003]. The degree distribution affects the resilience of complex food webs against random failures and intentional attacks: food webs showing right-skewed distributions, like the one described in this study, are more vulnerable to the removal of the most connected species or hubs, with the potential of producing secondary extinctions and a catastrophic fragmentation of the network [@Albert2000; @Dunne2002; @Eklof2006].

It is suggested that the small-world pattern, i.e., a network with short path length and high clustering coefficient is not frequent in complex marine food webs, mainly due to a low clustering coefficient compared to random networks [@Dunne2002b; @Marina2018a]. However, the food web of the MPA N-BB does display a small-world pattern. Consequences of this could be of great importance in recognizing evolutionary paths and the vulnerability to perturbations [@Montoya2002]. On the one hand, a short path length implies a rapid spread of an impact (e.g., contaminant, population fluctuation, local extinction) throughout the network and, at the same time, a more potentially adaptive dynamics in face of external perturbations [@Montoya2002; @Williams2002]. On the other hand, a high clustering coefficient indicates the formation of subnetworks composed only by the neighbours of particular species. This translates into a greater resistance of the network due to the confinement of perturbations mainly within subnetworks and not spreading between them [@Kortsch2019; @Heer2020]. Overall, a small-world topology provides ecological networks with a greater resilience and resistance [@Bornatowski2017; @Dormann2017].

Omnivory acts as a buffer to changes as the ecosystem presents alternative energy pathways in the face of perturbations; i.e., reducing the risk of cascading extinctions following primary loss of species [@Borrvall2000]. This is based on the fact that omnivores are species able to adapt faster and to a wider range of environmental conditions by changing their foraging habits to feed on the most abundant prey [@Fagan1997]. Furthermore, omnivory can be analysed from the interaction point of view: theoretical studies have identified omnivorous interactions as a possible candidate for a keystone interaction, sensu @Kadoya2018, highlighting the importance of omnivory inn stabilizing food web dynamics [@McCann1997; @Neutel2002]. The high proportion of omnivory in the food web of the MPA N-BB suggests that the network might be robust to variations in prey abundances, which could increase food web's persistence and stability [@Stouffer2010].

To resume, the food web of the MPA Namuncurá - Banco Burdwood presents a combination of network properties that make it unique in terms of network resolution, complexity, and structural pattern. All this suggest that the ecosystem might be fragile to external perturbations targeting highly connected species. However, structural properties might provide resilience and resistance with the final outcome of a rearrange structure maintaining their functions.

@@

Describing species’ roles in food webs provides a toolbox with which to assess the significance of species in terms of community's functioning and overall stability [@Thebault2010; @Cirtwill2018]. We used a range of descriptors to characterise the dynamic and multifaceted nature of the species forming the MPA N-BB food web.
